---
title: "Assignment_7_leigh_garcia_luna"
author: "Kat Leigh"
date: "5/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sensitivity)

```

# Using Leslie Matrices to evolve a population of rare rabits in an urban park
- given initially 10 adults allowed to grow for 20 years

```{r}
source("evolve_pop.R")

# fertility rates
F1 = 0.0
F2 = 2
F3 = 6
F4 = 1

# survivability - based on mortality rates per thousand per decade
s1 = .8 
s2 = .85
s3 = .65
s4 = .1

# initial population parameters
ini = c(0,0,10,0)
nyears = 20
fert = c(F1, F2, F3, F4)
surv = c(s1, s2, s3, s4)
rabit_pop=evolve_pop(fert, surv, ini, nyears)

head(rabit_pop)

# graph differnt components of the output
# total population

# add decade 
yearz = seq(from=1, to=nyears)
rabit_tot = cbind.data.frame(yearz=yearz, poptot=rabit_pop$poptot)

# plot total population per year
ggplot(rabit_tot, aes(yearz, poptot))+
  geom_col()+
  labs(y="Total Popultation")
```

```{r}
# plot total population per year broken down into age classes
rabit_ages = cbind.data.frame(yearz=yearz, t(rabit_pop$popbyage))
rabit_agesl = rabit_ages %>% gather(key="agecat", value="pop",-yearz)
ggplot(rabit_agesl, aes(yearz, pop, fill=agecat))+
  geom_col()+
  labs(y="Population", fill="Age Group (years)")+
  scale_fill_discrete(labels = c("0-1", "1-2", "2-3", "3-4"))
```
# find key values: total population & number of young rabits at year 20
```{r}
# total population
tot_pop_20 <- round(rabit_tot$poptot[20],0)
tot_pop_20

# number of young rabits
young_20 <- rabit_agesl %>% 
  filter(agecat == "1",
         yearz == "20")

young_20_pop <- round(young_20$pop[1],0)  
young_20_pop
```
# how does the rabit population change if hawks are introduced?

- Hawks reduce the survivability of the young and sub-adults age classes to:
     - between 0.65 and 0.75 for ages 0-1 and
     - between 0.75 and 0.8 for Ages 1-2.
- distributions are uniform.

Perform a sensitivity analysis on the survivability of Young Age and Sub-Adults age class parameters. .  Use Sobel to generate variation in the survivability parameters.

```{r}

nsample=200

# create our two samples for Sobel
# first do our survivability
ss1 = cbind.data.frame(s1=runif(min=0.65, max=0.75, n=nsample), s2 = runif(min=0.75, max=0.8, n=nsample))

ss2 = cbind.data.frame(s1=runif(min=0.65, max=0.75, n=nsample), s2 = runif(min=0.75, max=0.8, n=nsample))

# get sobel samples
sens_hawkbits=soboljansen(model = NULL, ss1, ss2, nboot = 100)

head(sens_hawkbits$X)
nsim=nrow(sens_hawkbits$X)

```
# problems in this chunk. I thought it was the non-changing values being single numbers, so I created them as lists instead, but that didn't help.

```{r}
# run the model w/ the sobel values

ini = c(0, 0, 10, 0) # initial population for each age class
nyears = nyears

# non-changed age class values
s3 = rep(s3, 200)
s4 = rep(s4, 200)
F1 = rep(F1, 200)
F2 = rep(F2, 200)
F3 = rep(F3, 200)
F4 = rep(F4, 200)

# as before combine our application of the the dynamics model - for each
# parameter set, with code to extract our metric of interest (final population)
p_wrapper = function(s1=ss1, s2=ss2, s3=s3, s4=s4, F1=F1, F2=F2, F3=F3, F4=F4, use_func, initialpop, nstep) {
fertility=c(F1,F2,F3, F4)
survivability= c(s1,s2,s3,s4)
res = use_func(survivability = survivability, fertility = fertility, initialpop=initialpop, nstep=nstep)
# now return the final population total
return(totpop=res$poptot[nstep])
}

# use pmap here so we can specify rows of our sensitivity analysis parameter object 
res = as.data.frame(sens_hawkbits$X) %>% pmap_dbl(p_wrapper, initialpop=ini, nstep=nyears, use_func=evolve_pop)
         
# plot results (variation in final population across all parameter)
# ggplot needs a dataframe - so do a quick conversion with data.frame
ggplot(data.frame(finalpop=res), aes(x=finalpop))+geom_density()

```


Then for each parameter set, compute total rabbit population after 20 years.