---
title: "Rabbits & Hawks"
author: "Hannah Garcia, Kat Leigh, & Anthony Luna"
date: "5/15/2021"
output:
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r}
# Upload packages
library(sensitivity)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r}
# read in function
source("evolve_pop.r")
```

```{r}
# set up fertility and survivability of rabbits
# fertility rates
f1 = 0
f2 = 1
f3 = 3
f4 = 0.5

# survivability 
p1 = 0.8
p2 = 0.85
p3 = 0.65
p4 = 0.1

# set up initial population parameters
ini = c(0, 0, 10, 0)
nyears = 20
fert_rabbits = c(f1,f2,f3,f4)
surv_rabbits = c(p1, p2, p3, p4)
rabbits_pop=evolve_pop(fert_rabbits, surv_rabbits, ini, nyears)
  
head(rabbits_pop)

# q1: what is the total rabbit population after 20 years?
tot_pop_20 <- round(rabbits_pop$poptot[20],0)
tot_pop_20

## ---------------------------------------------------------------------------
# extra - plots to visualize rabbit population
# add year 
year = seq(from=1, to=nyears)
rabbits_tot = cbind.data.frame(year=year, poptot=rabbits_pop$poptot)
ggplot(rabbits_tot, aes(year, poptot))+geom_col()+labs(y="total population")

# plot information about ages
rabbits_ages = cbind.data.frame(year=year, t(rabbits_pop$popbyage))
rabbits_agesl = rabbits_ages %>% gather(key="agecat", value="pop",-year)
ggplot(rabbits_agesl, aes(year, pop, fill=agecat))+geom_col()+labs(y="population", fill="age group")

# q2: how many young rabbits (first age class) are there in the population at that time?
young_20 <- rabbits_agesl %>% 
  filter(agecat == "1",
         year == "20")

young_20_pop <- round(young_20$pop[1],0)  
young_20_pop

```



## sensitivity analysis

hawks generally only eat  younger rabbits- thus they reduce the survivability of the young and sub-adults age classes (the first two classes).  the estimates are that survivability reduced to between 0.65 and 0.75 for ages 0-1 and between 0.75 and 0.8 for ages 1-2. you can assume that distributions are uniform

plot how total rabbit population after 20 years varies with a) survivability of young age and b) young adult.  also generate a box plot of the variation in total rabbit population.  how does this compare with total rabbit population after 20 years in your original population model (without the hawk related change to survivability)?

```{r sobel-data-setup}
nsample = 200
# set up fertility and survivability of rabbits
# fertility rates
parm_sample_1 <- cbind.data.frame(
  # fertility
  f1 = 0,
  f2 = 1,
  f3 = 3,
  f4 = 0.5,
  # survivability 
  p1 = runif(min=0.65,max=0.75,n=nsample),
  p2 = runif(min=0.75,max=0.85,n=nsample),
  p3 = 0.65,
  p4 = 0.1
)

parm_sample_2 <- cbind.data.frame(
  # fertility
  f1 = 0,
  f2 = 1,
  f3 = 3,
  f4 = 0.5,
  # survivability 
  p1 = runif(min=0.65,max=0.75,n=nsample),
  p2 = runif(min=0.75,max=0.85,n=nsample),
  p3 = 0.65,
  p4 = 0.1
)

sensitivity_data <- soboljansen(model=NULL,parm_sample_1,parm_sample_2,nboot=100)

head(sensitivity_data$X)
nrow(sensitivity_data$X)
```

```{r ad-hoc-function}
p_wrapper_tot = function(p1, p2, p3, p4, f1, f2, f3, f4, use_func, initialpop, nstep) {
  fertility=c(f1,f2,f3,f4)
  survivability= c(p1, p2, p3, p4)
  res = use_func(survivability =survivability, fertility = fertility, initialpop=initialpop, nstep=nstep)
  # return the final population total
  return(finalpop=res$poptot[nstep])
}

```


```{r implement-sensitivity-analysis}

nyears=20
# use pmap here so we can specify rows of our sensitivity analysis parameter object 
res_tot = as.data.frame(sensitivity_data$X) %>% pmap_dbl(p_wrapper_tot, initialpop=ini, nstep=nyears, use_func=evolve_pop)


```





```{r}

# plot results (variation in population at year 20 across survivability of young age and young adult rabbits)
# ggplot needs a dataframe - so do a quick conversion with data.frame
ggplot(data.frame(finalpop=res_tot), aes(x=finalpop))+geom_density()

# most of the values result in populations less than 15,000 rabbits

# boxplot of the same data
ggplot(data.frame(finalpop=res_tot), aes(x="", y=finalpop/1000) )+geom_boxplot(fill="blue")+
  theme(axis.title.x = element_blank())+labs(y="Final Pop (in 1000s)")

# give our results to sensitivity structure

sens_rabbits=tell(sensitivity_data, res_tot)

# look at results
sens_rabbits$S
sens_rabbits$T
```
graph the most sensitive parameter (survivability of young rabbits S-value: `r sens_rabbits$S[5,1]`, T-value of: `r sens_rabbits$T[5,1]`)

```{r}
orig_8 <- data.frame(x=0.8,y=tot_pop_20)

tmp = cbind.data.frame(sens_rabbits$X, poptot=sens_rabbits$y)
ggplot()+
  geom_point(data=filter(tmp, p2> .845), aes(p1, poptot), color= "mediumseagreen", alpha=.5)+
  geom_point(data=filter(tmp, p2< .845), aes(p1, poptot), color= "black", alpha=.3)+
  stat_smooth(data=tmp, aes(p1, poptot), se=FALSE, size=1.5, alpha=.7, fullrange = TRUE)+
  labs(x="Survivability of rabbits aged 0-1",y="Total population after 20 years",
       caption= str_wrap("The dots depict the predicted total population of rabbits at the end of 20 years when the surviveability of ages 0-1 ranges between 0.65 and 0.75, and the survivability for ages 1-2 is between 0.75 and 0.85 so as to simulate the presence of hawk predation. The central tendancy and trend of the data is indicated by the blue line. The predicted rabbit population without hawks (survivability 0-1 = 0.8, 1-2 = 0.85) is also shown with a triangle. For comparison, all results with the survivability for ages 1-2 roughly equal to 0.85 are colored in seagreen, since this corresponds to the age 1-2 survivability of the no-hawk scenario. Compared to the no-hawk scenario, rabbit populations will be noteably smaller with hawks. In fact, the maximum population size possible with hawk predation is 21,997, which is roughly 40% smaller than without hawks.", 115))+
  scale_x_continuous(limits=c(.65,.81))+
  geom_point(data=orig_8, aes(x=x,y=y), color= 'mediumseagreen', fill= 'mediumseagreen', size=3, shape=17)

```


```{r}
# Kat's extra funzies young rabbit graph

p_wrapper_young = function(p1, p2, p3, p4, f1, f2, f3, f4, use_func, initialpop, nstep) {
  fertility=c(f1,f2,f3,f4)
  survivability= c(p1, p2, p3, p4)
  res = use_func(survivability =survivability, fertility = fertility, initialpop=initialpop, nstep=nstep)
  # return the young population total
  return(popbyage=res$popbyage[1,nstep])
}

res_young = as.data.frame(sensitivity_data$X) %>% pmap_dbl(p_wrapper_young, initialpop=ini, nstep=nyears, use_func=evolve_pop)

# give our young rabbit results to our sensitivity structure
sens_youngbits=tell(sensitivity_data, res_young)

orig_8 <- data.frame(x=0.8,y=young_20_pop)

tmp = cbind.data.frame(sens_youngbits$X, pop1=sens_youngbits$y)
ggplot()+
  geom_point(data=filter(tmp, p2> .845), aes(p1, pop1), color= "mediumseagreen", alpha=.5)+
  geom_point(data=filter(tmp, p2< .845), aes(p1, pop1), color= "black", alpha=.3)+
  stat_smooth(data=tmp, aes(p1, pop1), se=FALSE, size=1.5, alpha=.7, fullrange = TRUE)+
  labs(x="Survivability of rabbits aged 0-1",y="Young rabbit population after 20 years",
       caption=str_wrap("The dots depict the predicted total population of young rabbits at the end of 20 years when the surviveability of ages 0-1 ranges between 0.65 and 0.75, and the survivability for ages 1-2 is between 0.75 and 0.85 so as to simulate the presence of hawk predation. The central tendancy and trend of the data is indicated by the blue line. The predicted young rabbit population without hawks (survivability 0-1 = 0.8, 1-2 = 0.85) is also shown with a triangle. For comparison, all results with the survivability for ages 1-2 roughly equal to 0.85 are colored in seagreen, since this corresponds to the age 1-2 survivability of the no-hawk scenario. Compared to the no-hawk scenario, young rabbit populations will tend to be smaller with hawks.", 115))+
  scale_x_continuous(limits=c(.65,.81))+
  geom_point(data=orig_8, aes(x=x,y=y), color= 'mediumseagreen', fill= 'mediumseagreen', size=3, shape=17)
```
 